Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-23 paper/paper.tex
Edits to abstract and introduction

\begin{abstract}
Abstract goes here.
\end{abstract}

@@

Scientific workflows, usually represented as directed acyclic graphs (DAG), are
the important class of applications that have been studied in the context of
resource management and task scheduling on grid and utility computing systems.
However, large computational applications are often composed of several 
inter-related workflows grouped into {\em ensembles}. All the
workflows in an ensemble typically have a similar structure, but they differ by
input data, number of tasks and individual task sizes. A good example of
scientific workflow ensemble comes from the CyberShake
application~\cite{Callaghan11}, which calculates seismic hazards for a given
geographical region such as California. In order to produce a hazard map, the
hazard curves need to be computed for a set of geographical sites, while each
site requires running a single complex scientific workflow. Workflows for each
site may differ not only in their parameters, but also in their priority; e.g. some of
the points on the map correspond to urban areas or strategic locations such as
power plants, whereas others may be located in less populated regions. Another
example of workflow ensembles comes from astronomical applications, such as
Montage~\cite{Deelman08} or search for exo-solar planets from Kepler project
data~\cite{vockler11}, where multiple, different workflows are used to
process data for different areas of the sky.

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
computational resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost, on-demand computing resources~\cite{Ostermann10,Keahey09}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give users more flexibility to creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the dynamically changing
computing demands of the application, often called auto-scaling. However, giving users 
more control over the resources also requires the development of new methods for task 
scheduling and resource provisioning. The resource management decisions required 
in cloud scenarios not only have to take into account traditional performance-related 
metrics such as the workflow makespan or the resource utilization, but must also now consider
budget constraints, since the resources from public (commercial) cloud providers 
usually have monetary costs associated with them~\cite{Durkee10}.

In this paper, we aim to gain insight into resource management challenges when executing 
scientific workflow ensembles on clouds. In particular we address a new important problem of maximizing
the number of completed workflows from an ensemble under both budget and
deadline constraints. The motivation is to answer the fundamental question of a
researcher: How much computation can be completed given the limited budget and 
timeframe of a research project? The goals of this paper are to discuss and assess possible
static and dynamic strategies for both task scheduling and resource
provisioning. Based on the knowledge of workflow scheduling algorithms we
analyze strategies for on-line scheduling of individual tasks as well as static
algorithms that rely on the information about the workflow structure (critical
paths and workflow levels) and estimates of task runtimes. In addition, we
analyze a hybrid workflow-aware dynamic scheduling algorithm, which uses the
workflow structure information to estimate which workflows should be rejected
from the ensemble due to the constraints. We evaluate the algorithms via simulation. 
We have developed cloud workflow simulator based on
CloudSim~\cite{Calheiros11}, which models the infrastructure and the application. The
algorithms are subject to evaluation on a set of scientific workflow ensembles with a broad range of
budget and deadline parameters.

The paper is organized as follows: after analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} gives the details about the
infrastructure and application model we are targeting in this paper.
In Section~\ref{sec:algorithms} we describe the dynamic and
static algorithms we developed. Section~\ref{sec:performance} presents test
scenarios, metrics and results of simulation studies of these algorithms.
Finally, general conclusions, lessons learned and future work are outlined in
Section~\ref{sec:conclusions}.

>>>>

\begin{abstract}

Large-scale computational applications represented as scientific workflows, are
often grouped into ensembles composed of several inter-related workflows. In
this paper, we address a new and important problem concerning the efficient 
management of such ensembles under both budget and deadline constraints on 
Infrastructure-as-a-Service (IaaS) clouds. IaaS clouds are characterized by
on-demand resource provisioning capabilities and a pay-per-use model.  We discuss,
develop, and assess algorithms based on static and dynamic strategies for both
task scheduling and resource provisioning. We assume that the workflows within
an ensemble are prioritized by the user. We evaluate the algorithms via
simulation using a set of scientific workflow ensembles with a broad range of
budget and deadline parameters, identifying which algorithm performs best for 
different application scenarios and constraint ranges. We found that the key 
factor determining the performance of an algorithm is its ability to decide which
workflows in an ensemble to admit or reject for execution. The results
show that the admission procedure based on workflow structure and estimates of
task runtimes can significantly improve the quality of solutions.

\end{abstract}

@@

Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
an important class of distributed applications that have been studied in the context of
resource management and task scheduling on grid and utility computing systems.
Workflow applications for large computational problems are often composed of 
several inter-related workflows grouped into {\em ensembles}. All the
workflows in an ensemble typically have a similar structure, but they differ in their
input data, number of tasks, and individual task sizes. The CyberShake 
application~\cite{Callaghan11} is a good example of an application that requires 
scientific workflow ensembles. CyberShake computes seismic hazard curves for a given
geographical location. In order to produce a hazard map for a region such as 
Southern California, many hazard curves need to be computed for a large number of 
geographical sites, where each site requires running a single, complex scientific 
workflow. Workflows for individual sites may differ not only in their parameters, 
but also in their priority. For example, some sites may be heavily populated areas 
or in strategic locations such as power plants, while others may be less important. 
Other examples of workflow ensembles come from astronomical applications that create
science-grade image mosaics~\cite{Deelman08} or search for extrasolar 
planets~\cite{vockler11}, where multiple, different workflows are used to
process data for different areas of the sky.

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
computational resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost, on-demand computing resources~\cite{Ostermann10,Keahey09}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give users more flexibility in creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the dynamically changing
computing demands of the application, often called auto-scaling. However, giving users 
more control over the resources also requires the development of new methods for task 
scheduling and resource provisioning. The resource management decisions required 
in cloud scenarios not only have to take into account traditional performance-related 
metrics such as workflow makespan or resource utilization, but must also now consider
budget constraints, since the resources from public (commercial) cloud providers 
usually have monetary costs associated with them~\cite{Durkee10}.

In this paper, we aim to gain insight into resource management challenges when executing 
scientific workflow ensembles on clouds. In particular, we address a new and important 
problem of maximizing the number of completed workflows from an ensemble under both budget and
deadline constraints. The motivation for this work is to answer the fundamental question of 
concern to a researcher: How much computation can be completed given the limited budget and 
timeframe of a research project? The goals of this paper are to discuss and assess possible
static and dynamic strategies for both task scheduling and resource
provisioning. Based on the knowledge of workflow scheduling algorithms we
analyze strategies for on-line scheduling of individual tasks as well as static
algorithms that rely on the information about the workflow structure (critical
paths and workflow levels) and estimates of task runtimes. In addition, we
analyze a hybrid workflow-aware dynamic scheduling algorithm, which uses the
workflow structure information to estimate which workflows should be rejected
from the ensemble based on the constraints. We evaluate these algorithms using a
simulator we have developed using CloudSim~\cite{Calheiros11}, which models the 
infrastructure and the application. The algorithms were evaluated on a set 
of scientific workflow ensembles from several different applications using a 
broad range of budget and deadline parameters.

The main contributions of this paper are:
\begin{itemize}
  \item a definition of the problem of scheduling prioritized workflows grouped into
  ensembles, under both budget and deadline constraints,
  \item an analysis of static and dynamic strategies for task scheduling and
  resource provisioning on clouds, together with admission procedures to
  predict which workflows from the ensemble can be completed,
  \item the development of several static and dynamic algorithms and of a simulator based
  on the defined application and infrastructure model,
  \item a performance evaluation of the algorithms on a set of synthetic benchmark
  workflow ensembles based on important, real scientific applications,
  \item an analysis of performance results that assesses the advantages and
  disadvantages of the proposed algorithms on different application scenarios and
  varying constraint values.
\end{itemize}

The paper is organized as follows: after an analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
infrastructure and application model we are targeting in this paper.
In Section~\ref{sec:algorithms} we describe the dynamic and
static algorithms we developed. Section~\ref{sec:performance} presents test
scenarios, performance metrics, and the results of simulation studies of these algorithms.
Finally, general conclusions, lessons learned, and future work are outlined in
Section~\ref{sec:conclusions}.
